---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# diemr

<!-- badges: start -->
[![CRAN version](https://www.r-pkg.org/badges/version/diemr)](https://CRAN.R-project.org/package=diemr)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/diemr)](https://cran.r-project.org/package=diemr)
[![Monthly downloads](https://cranlogs.r-pkg.org/badges/last-month/diemr)](https://cran.r-project.org/package=diemr)
<!-- badges: end -->

_diemr_ incorporates the diagnostic index expectation maximisation algorithm used to estimate which genomic alleles belong to either side of a barrier to geneflow. 

## Installation

To start using _diemr_, load the package or install it from CRAN if it is not yet available:

``` {r, eval = FALSE}
if(!require("diemr", character.only = TRUE)){
    install.packages("diemr", dependencies = TRUE)
    library("diemr", character.only = TRUE)
}
# Loading required package: diemr
```

Set working directory to a location with read and write privileges.

## Check data format

Next, assemble paths to all files containing the data to be used by _diemr_. Here, we will use a tiny example dataset for illustration that is included in the package. A good practice is to check that all files contain data in correct format for all individuals and markers.

```{r, eval = FALSE}
filepaths <- system.file("extdata", "data7x3.txt",
                         package = "diemr")
CheckDiemFormat(filepaths, ploidy = list(rep(2, 6)), ChosenInds = 1:6)
# File check passed: TRUE
# Ploidy check passed: TRUE
```

If the `CheckDiemFormat()` function fails, work through the error messages and fix the stored input files accordingly. The algorithm repeatedly accesses data from the harddisk, so seeing the passed file check prior to analysis is critical.

```{r, eval = FALSE}
diem.res <- diem(files = filepaths,
				 ploidy = list(rep(2, 6)), 
				 ChosenInds = 1:6,
                 verbose = TRUE,
                 nCores = 1)
```

The results including marker polarisation, marker diagnostic index and its support will be included in the list element `diem.res$DI`. Additional elements in the results list contain basic tracking information about the expectation maximisation iterations, with details provided in the folder _diagnostics_ in the working directory. 

